---
title: "Macro-Finance"
author: "Joan Antoni Segui"
output:
  html_document:
    toc: true
    toc_depth: 1
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'macro_fin.html'))})
---

# Problem Set 1


```{r }
library(haven)
library(data.table)
library(ggplot2)
library(reshape2)
JSTdatasetR4 <- as.data.table(read_dta("C:/Users/Jt_an/Google Drive/GSE - Finance/3 trim_/Macro Finance/Homework/JSTdatasetR4.dta"))
```

```{r }
unique(JSTdatasetR4$country)
```

```{r }
JSTdatasetR4[ country=="Spain" & crisisJST==1, ]
```

```{r warning=FALSE}
data = JSTdatasetR4[country=="Spain" & year >= 1972 & year <=1982, .(year, housing_capgain,eq_capgain, inflation= (diff(cpi)/cpi))]

real_eq_capgains = (1+data$eq_capgain)/(1+data$inflation)-1
real_housing_capgains = (1+data$housing_capgain)/(1+data$inflation)-1

index_real_eq = 1
index_real_housing= 1
for (i in 2:11){
  index_real_eq[i]=index_real_eq[i-1]*(1+ real_eq_capgains[i])
  index_real_housing[i]=index_real_housing[i-1]*(1+ real_housing_capgains[i])
}



data_plot=melt(as.data.table(cbind(year=data$year,"Real Stock Price" = index_real_eq,"Real House Price"=index_real_housing)),id.vars = "year")


ggplot(data_plot, aes(year,value,colour=variable))+
  geom_line()+
  labs(title = "1977 Spanish Crisis ", x = "Year", y = "Index. 1977=1", color = "")+
  geom_vline(xintercept =1977)+
  scale_x_continuous(breaks=data_plot$year)+ 
  annotate("text", x = 1978.5, y = 1, label = "Start of the Crisis")

```


# Problem Set 2



```{r }
library(data.table)
library(ggplot2)
library(kableExtra)



```


```{r }
SP500<- as.data.table(read.csv("C:/Users/Jt_an/Google Drive/GSE - Finance/3 trim_/Macro Finance/Homework/HW2/^GSPC.csv"))
ICSA <- read.csv("C:/Users/Jt_an/Google Drive/GSE - Finance/3 trim_/Macro Finance/Homework/HW2/ICSA.csv")


SP500 <- SP500[,Date := as.Date(Date)]


SP500 <- SP500[,weekly:=(Adj.Close-shift(Adj.Close,5))/shift(Adj.Close,5)][,monthly:=(Adj.Close-shift(Adj.Close,22))/shift(Adj.Close,22)]

SP_weekly<-head(SP500[order(weekly)],20)
SP_monthly<-head(SP500[order(monthly)],20)

SP_monthly[,id:=paste0(year(Date),"-",month(Date))]
list<-SP_monthly[,unique(id)]


tablemonthly<- cbind(SP_monthly[1:12, .(return=head(monthly,1)), by=id],claims=c("NA","NA",10667000,"NA","NA"))

tableweekly<- cbind(SP_weekly[Date == "1929-10-29" | Date == "1987-10-19" | Date == "1932-09-14" | Date == "1932-10-10" | Date == "1933-07-21",
          .(Date, return=weekly)],claims= c("NA",301000,"NA","NA","NA"))
```

```{r }
k1<-kable(tablemonthly, caption = "Top-5 monthly falls + initial claims in that month")
kable_styling(k1, latex_options = "hold_position")
k2<-kable(tableweekly, caption = "Top-5 weekly falls + initial claims in that week")
kable_styling(k2, latex_options = "hold_position")


```

# Problem Set 4


```{r }
library(data.table)
library(haven)
library(kableExtra)
```



```{r }
JSTdatasetR4 <- as.data.table(read_dta("C:/Users/Jt_an/Google Drive/GSE - Finance/3 trim_/Macro Finance/Homework/JSTdatasetR4.dta"))
```


```{r }

JSTdatasetR4<-JSTdatasetR4[country=="Spain" & between(year,1900,2015 ) | country=="Switzerland" & between(year,1900,2015 ), .(year, country,inflation= ((cpi-shift(cpi))/shift(cpi)),log_rconpc = log(rconpc)-shift(log(rconpc)),eq_tr, cpi, bill_rate, rconpc), by="country"][!is.na(inflation)][,real_er := (1+eq_tr)/(1+inflation)-1][,real_sr := (1+bill_rate)/(1+inflation)-1][,equity_p:=real_er-real_sr]

JSTdatasetR4<- na.omit(JSTdatasetR4)

```

```{r}
table<-JSTdatasetR4[, .(risk_premium.lvl=mean(equity_p),risk_premium.sd=sd(equity_p),safe_rate.lvl=mean(real_sr), safe_rate.sd=sd(real_sr),corr_risky_cons=cor(log_rconpc,real_er), rconsg.lvl=mean(log_rconpc),rconsg.sd=sd(log_rconpc)) ,by="country"]


colnames(table) <- c("Country", "RPremium Level", "RPremium Sd", "Safe Rate Level", "Safe Rate Sd", "Corr(Risky Return, Consumption Growth","Consumption Growth Level","Consumption Growth Sd")
```


```{r }

k1<-kable(table, caption = "Spain and Switzerland comparison")
kable_styling(k1, latex_options = c("hold_position", "scale_down"))
```


# Problem Set 5


```{r }
library(data.table)
library(haven)
library(kableExtra)
library(xts)
JSTdatasetR4 <- as.data.table(read_dta("C:/Users/Jt_an/Google Drive/GSE - Finance/3 trim_/Macro Finance/Homework/JSTdatasetR4.dta"))
```


```{r }


p_ind<-1
esp<-JSTdatasetR4[country=="Spain" & !is.na(eq_capgain), .(year,eq_capgain,eq_dp,cpi,eq_tr,inflation= (cpi-shift(cpi))/shift(cpi))][,real_er:= (1+eq_tr)/(1+inflation)-1]


for (i in 1:(length(esp$year)-1)) {
  p_ind[i+1]<-p_ind[i]*(1+esp$eq_capgain[i+1])
}

esp<-cbind(esp,p_ind)

esp<-esp[,div_ind:=p_ind*eq_dp][,real_div_ind:=div_ind/cpi][,real_div_g := (real_div_ind-shift(real_div_ind)/shift(real_div_ind))][!is.na(real_div_g)][,lag.dp:=shift(eq_dp)][,log.r:=(log(1+real_er))][,log.g:=log(1+real_div_g)][,log.dp:=log(lag.dp)][!is.infinite(log.r) & !is.infinite(log.g) & !is.infinite(log.dp),]
```



```{r }
reg_r.dp <- lm(real_er ~ lag.dp, data=esp)
reg_g.dp <- lm(real_div_g ~ lag.dp, data=esp)

logreg_r.dp <- lm(log.r ~ log.dp, data=esp)
logreg_g.dp <- lm(log.g ~ log.dp, data=esp)
```



```{r }
ret_coef <- c(summary(reg_r.dp)$coefficients[2,1:4],summary(reg_r.dp)$r.squared)
g_coef <- c(summary(reg_g.dp)$coefficients[2,1:4],summary(reg_g.dp)$r.squared)
logret_coef <- c(summary(logreg_r.dp)$coefficients[2,1:4],summary(logreg_r.dp)$r.squared)
logg_coef <- c(summary(logreg_g.dp)$coefficients[2,1:4],summary(logreg_g.dp)$r.squared)

table<- rbind(ret_coef ,g_coef,logret_coef,logg_coef)
colnames(table)=c("Estimate","Std. Error","t stat","p-value", "R-squared")
rownames(table)=c("returns","dividend growth","log returns","log dividend growth")
k1<-kable(table, caption = "Spain Forecasting Regressions")
kable_styling(k1, latex_options = c("hold_position"))
```

Spanish returns seems predictable, the equity dividend yield coefficient is statistically significant at the .05 level. While for dividend growth it is not statistically significant. The coefficients are positive, so a higher equity dividend yield predicts a higher return and a higher dividend growth.

The same for the log regressions, for returns is significant at the 5% level,and not for dividend growth. A 1% increase in the equity dividend yield, increases returns by 0.15% and by 0.22% the dividend growth (now is significant at the 0.1 level).  

# Problem Set 6



```{r }
rm(list=ls())
library(haven)
library(data.table)
library(plm)
library(kableExtra)
```



```{r }

data <- as.data.table(read_dta("C:/Users/Jt_an/Google Drive/GSE - Finance/3 trim_/Macro Finance/Homework/JSTdatasetR4.dta"))
data <- data[year>=1870 & year<=2015, CumGDPGrowth_t:=(rgdppc-shift(rgdppc))/shift(rgdppc),by="country"][,CumGDPGrowth_t1 := (shift(rgdppc,type="lead")-shift(rgdppc))/shift(rgdppc),by="country"][,CumGDPGrowth_t2 := (shift(rgdppc,n=2,type="lead")-shift(rgdppc))/shift(rgdppc)][,gdpGrowth := (shift(rgdppc,n=1)-shift(rgdppc,n = 2))/shift(rgdppc,n = 2), by="country"][,inflation := (shift(cpi)-shift(cpi,n = 2))/shift(cpi,n = 2),by="country"][,LTinterestRate := shift(ltrate),by="country"][,STinterestRate := shift(stir),by="country"][,CAtoGDP := shift(ca)/shift(gdp),by="country"][,invToGDP := shift(iy),by="country"][,rcred := tloans/cpi,by="country"][,lag_grcred := (shift(rcred)-shift(rcred,n = 2))/shift(rcred,n=2),by="country"][,bigCrisis := ifelse(crisisJST==1 & lag_grcred>0.02,1,0),by="country"]
```


```{r }

regression1 <- plm(CumGDPGrowth_t ~ crisisJST + gdpGrowth + inflation + LTinterestRate + STinterestRate + CAtoGDP + invToGDP- 1, model="within",index="country",data=data, na.action=na.omit)


regression2 <- plm(CumGDPGrowth_t1 ~ crisisJST + gdpGrowth + inflation + LTinterestRate + STinterestRate + CAtoGDP + invToGDP - 1, model="within",index="country",data=data, na.action=na.omit)


regression3 <- plm(CumGDPGrowth_t2 ~ crisisJST + gdpGrowth + inflation + LTinterestRate + STinterestRate + CAtoGDP + invToGDP -1, model="within",index="country",data=data, na.action=na.omit)

```


```{r }
table1<-rbind(summary(regression1)$coefficients[1,],
summary(regression2)$coefficients[1,],
summary(regression3)$coefficients[1,])

rownames(table1)<-c("theta_0","theta_1","theta_2")

k1<-kable(table1, caption = "Crisis at different horizons")
kable_styling(k1, latex_options = "hold_position")
```


```{r }

regression4 <- plm(CumGDPGrowth_t ~ bigCrisis + gdpGrowth + inflation + LTinterestRate + STinterestRate + CAtoGDP + invToGDP  - 1, model="within",index="country",data=data, na.action=na.omit)


regression5 <- plm(CumGDPGrowth_t1 ~ bigCrisis + gdpGrowth + inflation + LTinterestRate + STinterestRate + CAtoGDP + invToGDP - 1, model="within",index="country",data=data, na.action=na.omit)

regression6 <- plm(CumGDPGrowth_t2 ~ bigCrisis + gdpGrowth + inflation + LTinterestRate + STinterestRate + CAtoGDP + invToGDP - 1, model="within",index="country",data=data, na.action=na.omit)


```



```{r }
table2<-rbind(summary(regression4)$coefficients[1,],
summary(regression5)$coefficients[1,],
summary(regression6)$coefficients[1,])

rownames(table2)<-c("theta_0","theta_1","theta_2")

k2<-kable(table2, caption = "Crisis with real credit growth at different horizons")
kable_styling(k2, latex_options = "hold_position")
```


Financial crisis are costly and persistent. Cumulative real GDP growth at t-1 to t is on average 1.74 ppts below the trend. From t-1 to t+1 is below 4.75 ppts. From t-1 to t+2 is below 4.9. If we define crisis as in the first scenario with the additional constraint the lagged credit growth being bigger than the mean (+0.02), we get similar results.








































